Design, Implementation, and Evaluation of Open Power Quality

### 2020-05-15 01-abstract.tex
update

\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture which must accomodate distributed, intermittent generation.  This transition also means that the formerly sharp distinction between energy producers (i.e. utility companies) and consumers (residences, businesses, etc) are blurring: residences can now both produce and consume energy, making energy policy more complex. The Open Power Quality (OPQ) project began in 2013 with the goal of designing and implementing a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.  Since then, we have designed low cost hardware devices that monitor power quality  and low-cost cloud-based software services that can economically analyze the data and detect a variety of anomalies. In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.  Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions. We conclude that OPQ provides a new and useful approach to power quality monitoring.}

>>>>

\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture with significant distributed, intermittent generation.  This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) will blur: customers will both produce and consume energy, making energy management and public policy more complex. The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.  In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.  Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions. We conclude that OPQ provides a new and useful approach to power quality monitoring.}

### 2020-05-15 02-intro.tex
update

Power quality is not currently a concern for most people in developed nations. Just like most people in developed nations assume that their tap water is of adequate quality to drink, most also assume that their electricity is of adequate quality to power their homes and appliances without causing harm. And, in both cases, most people assume that public utilities will monitor and correct any quality problems if they occur.

Successfully maintaining adequate power quality and providing sufficient amounts of it to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years. In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing public concern and interest. First, there is a global need to shift to renewable energy sources. Second, modern consumer electronics place more stringent demands on power quality.  Third, effective policy making in this modern context requires better public access to power quality data.

{\em We need more renewable energy.} There is now a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar. But the economic, environmental, and political advantages of renewable energy comes with significant new technical challenges. Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed). In addition, renewable energy generation can be distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).

One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging. This problem is inversely proportional to the size of the electrical grid. For example, in Hawaii on the island of Oahu, the public demand for rooftop solar exceeds Hawaiian Electric Company's ability to utilize it and maintain adequate power quality. As a result, in many neighborhoods, new rooftop solar installation is prohibited because Hawaiian Electric fears its impact on not only local power quality but global grid stability \cite{TBD}.

{\em Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality. Only a few decades ago, computers were a rare presence outside of labs and large institutions. Today, computers are everywhere: embedded in phones, washers, refrigerators, thermostats, and so forth. These electronic devices not only have higher power quality requirements, but some of them actually introduce power quality problems in the form of harmonic distortion. Poor power quality can result in electronic devices failing unpredictably, and/or decreasing their lifespan.

>>>>

Power quality is not currently a concern for most people in developed nations. Just like most people in developed nations assume that their tap water is of adequate quality to drink, most also assume that their electricity is of adequate quality to power their homes and appliances without causing harm. And, in both cases, most people assume that public utilities will appropriately monitor and correct any quality problems if they occur.

Successfully maintaining adequate power quality and providing sufficient amounts of it to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years. In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing concern and interest. First, there is a global need to shift to renewable energy sources. Second, traditional top-down grid power quality monitoring may not be well suited to bottom-up, distributed generation associated with renewable energy sources such as roof-top solar. Third, modern consumer electronics place more stringent demands on power quality.  Finally, effective public policy making in this modern context can be aided by public access to power quality data.  Let's look at each of these in a bit more detail.

{\em 1. We need more renewable energy.} Concerns including pollution, environmental degradation, and climate change have produced a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar. But the economic, environmental, and political advantages of renewable energy comes with significant new technical challenges. Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed). In addition, renewable energy generation is often distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).

One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging. This problem is inversely proportional to the size of the electrical grid. For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.  As a result, in many neighborhoods, permits for new rooftop solar installation require complex and expensive "interconnection studies" because Hawaiian Electric fears the impact on grid stability \cite{trabish_solar_2014,anastasi_energy_2009}.

{\em 2. Traditional top-down monitoring is not well-suited to bottom-up energy generation.}

\cite{nakafuji_back--basics_2011}

{\em 2. Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality. Only a few decades ago, computers were a rare presence outside of labs and large institutions. Today, computers are everywhere: embedded in phones, washers, refrigerators, thermostats, and so forth. These electronic devices not only have higher power quality requirements, but some of them actually introduce power quality problems in the form of harmonic distortion. Poor power quality can result in electronic devices failing unpredictably, and/or decreasing their lifespan.
